---
title: Day-Zero Preparation
date: 2019-03-16
---


```{r echo=FALSE}
# please install_github("ijlyttle/vembedr")
library("vembedr")
```

We're all looking forward to getting together in Des Moines; we want everyone to be ready! This material may be useful to you regardless of your day-zero attendance.

We're going to have only a short time together in person, so we want to do all we can to help everyone get their computers set up to make it easier for us to collaborate. If you are experienced with R, git, and GitHub, and you are happy with your set-up, you need not do anything.

We plan to do a little bit each week until the Unconf starts.

Just about all of this material was compiled by Dr. Jenny Bryan of RStudio, in her book [Happy Git with R](https://happygitwithr.com). In our material, we will hit the high points, then link to the source material in case you want more background.

## Week of March 17

This week, we focus on a couple of tools:

- RStudio IDE [[Happy-Git ref.](https://happygitwithr.com/install-r-rstudio.html)]
- GitHub account [[Happy-Git ref.](https://happygitwithr.com/github-acct.html)]

With the RStudio IDE, we recommend using the [Preview version](https://www.rstudio.com/products/rstudio/download/preview/). This way, the the buttons and tabs on your IDE will be similar to the buttons and tabs on our IDE.

If you want to get a head-start on next week's activity, you can make sure that your R version is at least 3.4.0:

For your GitHub account:

- if you do not have an account already, please [sign up for one](https://github.com/join). For our purposes, you will not need anything other than a free account. Please keep your login-credentials handy.
- if you *do* already have an account, please make sure you have access to your login-credentials, i.e. you can enter them into the GitHub login page.

As you might imagine, you will need your login-credentials going forward.

With each week's exercise, we plan to release a screencast video where we go through the steps ourselves.

```{r echo=FALSE}
embed_youtube("1tpkC0z-4YE") %>%
  use_align("center") %>%
  use_rounded()
```

Having watched the video, I want to emphasize that:

- if you have version of RStudio IDE 1.2 or greater, you are likely just fine - I am intstalling the newest version just to make the demonstration.
- the preview version available, if and when you make your installation, may well be later than the version shown in the video - this is OK.

If you want to pick-and-choose topics:

- Introduction: 0:00
- RStudio IDE, macOS: 1:45
- RStudio IDE, Windows: 4:48
- GitHub account: 7:29
- Github issues: 8:33
- Summary: 10:46

### Verification

When you get through these steps, you are OK for this week:

1. With your (new) version of the RStudio IDE, open the Terminal pane, type the command `date`, then enter. Your computer should return the current date or date-and-time. If you are on Windows, and are asked to enter a new date, just hit **enter**.

2. Having signed into GitHub, go to [this issue](https://github.com/uncoast-unconf/uu-2019-day-zero/issues/11) on our issues page, either to introduce yourself, or to add an emoji reaction. Once you have done this, you are very welcome (in fact, encouraged) to participate in some of the [other issue discussions](https://github.com/uncoast-unconf/uu-2019/issues) at the main Uncoast Unconf repository, or even start new ones.

If you run into any problems, and are an Uncoast Unconf participant, you can check, then create a new [GitHub issue](https://github.com/uncoast-unconf/uu-2019-day-zero/issues) at the day-zero repository, or email Ian Lyttle at ian [dot] lyttle [at] se [dot] com.
