---
title: Day-Zero Preparation
date: 2019-03-24
---


```{r echo=FALSE}
# please install_github("ijlyttle/vembedr")
library("vembedr")
```

We're all looking forward to getting together in Des Moines; we want everyone to be ready! This material may be useful to you regardless of your day-zero attendance.

We're going to have only a short time together in person, so we want to do all we can to help everyone get their computers set up to make it easier for us to collaborate. If you are experienced with R, git, and GitHub, and you are happy with your set-up, you need not do anything.

We plan to do a little bit each week until the Unconf starts:

- [Week of March 17](#week-00): RStudio IDE and GitHub account
- [Week of March 24](#week-01): R and git installation

Just about all of this material was compiled by Dr. Jenny Bryan of RStudio and coworkers, in their book [Happy Git and Guthub for the useR](https://happygitwithr.com). In our material, we will hit the high points, then link to the source material in case you want more background.

## Week of March 24 {#week-01}

Thanks everyone for participating last week - if you have not had a chance to refresh your RStudio IDE and GitHub account, we invite you to go through the material, [found below](#week-00).

This week, we focus on installing foundational tools on our computer. Even if you already have R (which should be most of us), it can be useful to check that you are up-to-date:

- R: [[Happy-Git ref.](https://happygitwithr.com/install-r-rstudio.html)], [[CRAN R download](https://cloud.r-project.org)]
- git: [[Happy-Git ref.](https://happygitwithr.com/install-git.html)], Windows only: [[Git for Windows](https://gitforwindows.org/)]

Installing R and git are a little different between Windows and macOS, so we provide a screencast for each. That being said, there were a few things couple of things I would like to emphasize from the videos:

- I have edited out a lot of the time needed for the installation process, otherwise we would have 30-minute videos! I have edited out a few of the "clicks", but I have shown explicitly any option you have to confirm is set.

- Sorry for forgetting to change my Mac's RStudio IDE from "dark mode"; I'll remember next time :)

- If you do have to reinstall your packages, this usually gets *me* 80% of the way there:

```r
install.packages(c("tidyverse", "rmarkdown", "shiny", "devtools"))
```

### Windows

```{r echo=FALSE}
embed_youtube("Sp0i-OEDRgA") %>%
  use_align("center") %>%
  use_rounded()
```

### macOS

```{r echo=FALSE}
embed_youtube("K-92HPDX0F8") %>%
  use_align("center") %>%
  use_rounded()
```

### Verification

When you get through these steps, you are OK for this week:

#### R version

From the RStudio IDE Console pane (the R prompt):

```{r}
R.version.string
```

If your version is at least 3.4, you're OK.

#### Git version

From the RStudio IDE **Terminal** pane (access to your computer itself):

```bash
git --version
```

Your response should look something like this:

Windows:

```
git version 2.21.0.windows.1
```

macOS:

```
git version 2.17.2 (Apple Git-113)
```

---

If you run into any problems, and are an Uncoast Unconf participant, you can check, then create a new [GitHub issue](https://github.com/uncoast-unconf/uu-2019-day-zero/issues) at the day-zero repository, or email Ian Lyttle at ian [dot] lyttle [at] se [dot] com.

## Week of March 17 {#week-00}

This week, we focus on a couple of tools:

- RStudio IDE [[Happy-Git ref.](https://happygitwithr.com/install-r-rstudio.html)]
- GitHub account [[Happy-Git ref.](https://happygitwithr.com/github-acct.html)]

With the RStudio IDE, we recommend using the [Preview version](https://www.rstudio.com/products/rstudio/download/preview/). This way, the the buttons and tabs on your IDE will be similar to the buttons and tabs on our IDE.

If you want to get a head-start on next week's activity, you can make sure that your R version is at least 3.4.0:

For your GitHub account:

- if you do not have an account already, please [sign up for one](https://github.com/join). For our purposes, you will not need anything other than a free account. Please keep your login-credentials handy.
- if you *do* already have an account, please make sure you have access to your login-credentials, i.e. you can enter them into the GitHub login page.

As you might imagine, you will need your login-credentials going forward.

With each week's exercise, we plan to release a screencast video where we go through the steps ourselves.

```{r echo=FALSE}
embed_youtube("cKk7rioLFH0") %>%
  use_align("center") %>%
  use_rounded()
```

Special thanks to Sam Tyner and Amelia McNamara for providing feedback on drafts of this video; Amelia provided a script I was able to use verbatim: right at the start, where I'm asking for questions.

Having watched the video, I may have glossed over a few things, so I want to emphasize that:

- Winston and Binford are Basset Hounds, and will make noise as they please :)
- if you have version of RStudio IDE 1.2 or greater, you are likely just fine - I am installing the newest version just to make the demonstration.
- the preview version  that RStudio makes available, if and when you make your installation, may well be later than the version shown in the video - this is OK.

If you want to pick-and-choose topics:

- Introduction: 0:00
- RStudio IDE, macOS: 2:19
- RStudio IDE, Windows: 5:22
- GitHub account: 8:04
- GitHub issues: 9:08
- Summary: 11:21

### Verification

When you get through these steps, you are OK for this week:

1. With your (new) version of the RStudio IDE, open the Terminal pane, type the command `date`, then enter. Your computer should return the current date or date-and-time. If you are on Windows, and are asked to enter a new date, just hit **enter**.

2. Having signed into GitHub, go to [this issue](https://github.com/uncoast-unconf/uu-2019-day-zero/issues/11) on our issues page, either to introduce yourself, or to add an emoji reaction. Once you have done this, you are very welcome (in fact, encouraged) to participate in some of the [other issue discussions](https://github.com/uncoast-unconf/uu-2019/issues) at the main Uncoast Unconf repository, or even start new ones.

If you run into any problems, and are an Uncoast Unconf participant, you can check, then create a new [GitHub issue](https://github.com/uncoast-unconf/uu-2019-day-zero/issues) at the day-zero repository, or email Ian Lyttle at ian [dot] lyttle [at] se [dot] com.
